Data Flow Analysis\\in the Presence of Correlated Calls

### 2014-12-10 content/background.tex
comments

Similar to the IFDS algorithm, a practical implementation of the IDE algorithm requires
the input dataflow functions to be provided in their pointwise representation
as exploded supergraph edges labelled with micro-functions. Specifically,
the input is generally provided as a function $f: N^* \times D \times N^* \to (D \to F)$,
where $F$ is the set of representations of micro-functions from $L$ to $L$.
Given a (non-exploded) supergraph edge $n \to n'$ and a dataflow fact $d$,
$f(n,d,n')$ returns a map that gives for each fact $d'$ the micro-function $f$
that appears on the exploded supergraph edge $(n,d) \to (n',d')$.
Like in the IFDS algorithm, the function $f$ can be split up into separate functions that handle the cases when
the $n\to n'$ is an intraprocedural edge, a call-to-start edge, or an end-to-return edge.

>>>>

Similar to the IFDS algorithm, a practical implementation of the IDE algorithm requires
the input dataflow functions to be provided in their pointwise representation
as exploded supergraph edges labelled with micro-functions.
\mtodo{Should we cite the ``Practical extensions'' paper here that talks about the on-demand IFDS algorithm?}
Specifically,
the input is generally provided as a function $f: N^* \times D \times N^* \to (D \to F)$,
where $F$ is the set of representations of micro-functions from $L$ to $L$.
Given a (non-exploded) supergraph edge $n \to n'$ and a dataflow fact $d$,
$f(n,d,n')$ returns a map that gives for each fact $d'$ the micro-function $f$
that appears on the exploded supergraph edge $(n,d) \to (n',d')$.
Like in the IFDS algorithm, the function $f$ can be split up into separate functions that handle the cases when
the $n\to n'$ is an intraprocedural edge, a call-to-start edge, or an end-to-return edge.

### 2014-12-10 content/impl.tex
complete pass squeezing background on IDE algorithm
comments

\otodo{incorporate into previous section}
\subsection{Defining IFDS and IDE Problems}
%In Section~\ref{sec:bg}, we defined what IFDS and IDE problems are, their applications, and their constraints. In this section, we describe how to create instances of IFDS and IDE problems.

\commentout{
\subsubsection{Defining an IFDS Problem}\label{sec:ifdsdef}
Recall that an IFDS problem instance is defined as a five-tuple
\[
    (G^*,\ D,\ F,\ M_F,\,\sqcap)\,,
\]
where $G^*=(N^*,\,E^*)$ is the control-flow supergraph of the program, $D$ is the set of dataflow facts, $F\subseteq2^D\to2^D$ is a set of distributive dataflow functions, and the function $$M_F:\,E^*\to(2^D\to2^D)$$ maps the supergraph edges to dataflow functions, and is extended to paths by composition.

@@

The meaning of an edge in the exploded supergraph is the following.
Let $(n_1,\,d_1)$ and $(n_2,\,d_2)$ be two nodes in the exploded supergraph $G^\#$. Furthermore, assume that if fact $d_1$ at node~$n_1$ holds, then the fact~$d_2$ at node~$n_2$ also holds. Then there is an edge $(n_1,\,d_1),\,(n_2,\,d_2)\in E^\#$.
}

@@

To summarize, an IDE problem can be defined by a labeled exploded supergraph 
\begin{equation}(G^\#,\,L,\,\mathsf{EdgeFn})\,,\end{equation}
where each edge of the exploded supergraph corresponds to a micro function.

@@

%\paragraph{Converting IFDS problems to IDE problems}
%Since IDE is a generalization of IFDS, any IFDS problem can be converted into an equivalent IDE problem~\cite{sagiv1996precise}.
The lattice
for the equivalence transformation $\transEq$ 
is the two-point lattice
    $L^\equiv=\{\bot,\ \top\}$,
where $\bot$ means ``reachable'', and $\top$ means ``not reachable''. 
The edge functions $\edgefn^\equiv$ are defined as
\begin{equation}
    \edgefn^\equiv=
    \begin{cases}
      \lambda e\,.\,\lambda m\,.\,\bot&\text{if $d_1(e)=\Lambda$ and $d_2(e)\ne\Lambda$;}\\
      \lambda e\,.\,\id&\text{otherwise,}
    \end{cases}
\end{equation}
where $d_1(e)$ is the source fact of an edge $e$ and $d_2(e)$ is its target fact. 
At a ``diagonal'' edge from a $\Lambda$-fact to a non-$\Lambda$-fact
$d$, the micro function returns $\bot$
to make
the fact $d$ reachable. All other micro-functions are the identity.
The equivalence transformation is thus defined as:
  $\transEq((G^\#))=(G^\#,\,L^\equiv,\,\mathsf{EdgeFn}^\equiv)$.

@@

%\paragraph{Lattice Elements}
In the correlated-calls transformation, the 
%Just like in the equivalence transformation $\transEq$, the exploded supergraph for $\transCC$ is the same as in the original IFDS problem. The 
lattice elements are maps from receivers to sets of types:
    $\lcc=\left\{\,m:\ R\to2^T\right\}$,
where $R$ is the set of receivers and $T$ is the set of all types.
For each receiver $r$, the map gives an overapproximation of the possible
runtime types of $r$.
Sets of types are ordered by the superset relation, and this is lifted to maps from receivers
to sets of types, so the bottom element $\botcc$ maps every receiver to the set of all types,
and the top element $\topcc$ maps every receiver to the empty set of types.
During an actual execution, every receiver $r$ points to an object of some
runtime type. \otodo{What if $r$ points to null?} Therefore, a data flow fact
is unreachable along any feasible path if its corresponding lattice element
maps any receiver to the empty set of types.

@@

For each program point $n$, the result of an IFDS analysis gives
a set of facts $d$ that may be reached at $n$. The result of an IDE analysis
pairs each such fact $d$ with a lattice element $\ell$. Formally,
for an IFDS problem $P$, the result $\result_{\text{IFDS}}$ has type
$N \to D$. Similarly, for an IFDS problem $Q$, the result $\result_{\text{IDE}}$
has type $N \to D \times L$.

@@

\otodo{The normalization function $\mathcal N$ was only needed in composition. I therefore
    ``inlined'' it into composition to save space. This makes composition a bit harder to intuitively
    verify, but saves complexity here. The verification can be in the proof in the supplementary appendix.}
The composition of two micro-function representations is defined as follows:
$\setpair{I}{U} \circ \setpair{I'}{U'} = \setpair{\lambda r\,.\,(I(r) \cap I'(r))\cup U(r)}{\lambda r\,.\,(I(r)\cap U'(r))\cup U(r)}$.
The following lemma shows that this implementation corresponds to the composition
of the denoted functions:
\begin{lemma}\label{lem:equal}
    For any pair of micro-function representations $\setpair{I}{U}$, $\setpair{I'}{U'}$,
    \[
        \denote{\setpair{I}{U}} \circ \denote{\setpair{I'}{U'}}
        = 
        \denote{\setpair{I}{U}} \circ \denote{\setpair{I'}{U'}}
    \]
\end{lemma}
\otodo{Prove this lemma in the appendix. The parts of the proof can be found in section 5.}

>>>>

\commentout{
\otodo{incorporate into previous section}
\subsection{Defining IFDS and IDE Problems}
%In Section~\ref{sec:bg}, we defined what IFDS and IDE problems are, their applications, and their constraints. In this section, we describe how to create instances of IFDS and IDE problems.

\subsubsection{Defining an IFDS Problem}\label{sec:ifdsdef}
Recall that an IFDS problem instance is defined as a five-tuple
\[
    (G^*,\ D,\ F,\ M_F,\,\sqcap)\,,
\]
where $G^*=(N^*,\,E^*)$ is the control-flow supergraph of the program, $D$ is the set of dataflow facts, $F\subseteq2^D\to2^D$ is a set of distributive dataflow functions, and the function $$M_F:\,E^*\to(2^D\to2^D)$$ maps the supergraph edges to dataflow functions, and is extended to paths by composition.

@@

The meaning of an edge in the exploded supergraph is the following.
Let $(n_1,\,d_1)$ and $(n_2,\,d_2)$ be two nodes in the exploded supergraph $G^\#$. Furthermore, assume that if fact $d_1$ at node~$n_1$ holds, then the fact~$d_2$ at node~$n_2$ also holds. Then there is an edge $(n_1,\,d_1),\,(n_2,\,d_2)\in E^\#$.

@@

To summarize, an IDE problem can be defined by a labeled exploded supergraph 
\begin{equation}(G^\#,\,L,\,\mathsf{EdgeFn})\,,\end{equation}
where each edge of the exploded supergraph corresponds to a micro function.
}

@@

%\paragraph{Converting IFDS problems to IDE problems}
%Since IDE is a generalization of IFDS, any IFDS problem can be converted into an equivalent IDE problem~\cite{sagiv1996precise}.
The lattice
for the equivalence transformation $\transEq$ 
is the two-point lattice
    $L^\equiv=\{\bot,\ \top\}$,
where $\bot$ means ``reachable'', and $\top$ means ``not reachable''. 
The edge functions $\edgefn^\equiv$ are defined as
\begin{equation}
    \edgefn^\equiv=
    \begin{cases}
      \lambda e\,.\,\lambda m\,.\,\bot&\text{if $d_1(e)=\Lambda$ and $d_2(e)\ne\Lambda$;}\\
      \lambda e\,.\,\id&\text{otherwise,}
    \end{cases}
\end{equation}
where $d_1(e)$ is the source fact of an edge $e$ and $d_2(e)$ is its target fact. 
At a~``diagonal'' edge from a $\Lambda$-fact to a non-$\Lambda$-fact
$d$, the micro function returns $\bot$
to make
the fact $d$ reachable. All other micro-functions are the identity\mtodo{``the identity function'' or is it okay ot just say ``the identity''?}.
The equivalence transformation is thus defined as:
  $\transEq((G^\#))=(G^\#,\,L^\equiv,\,\mathsf{EdgeFn}^\equiv)$.

@@

%\paragraph{Lattice Elements}
In the correlated-calls transformation, the 
%Just like in the equivalence transformation $\transEq$, the exploded supergraph for $\transCC$ is the same as in the original IFDS problem. The 
lattice elements are maps from receivers to sets of types:
    $\lcc=\left\{\,m:\ R\to2^T\right\}$,
where $R$ is the set of receivers and $T$ is the set of all types.
For each receiver~$r$, the map gives an overapproximation of the possible
runtime types of $r$.
Sets of types are ordered by the superset relation, and this is lifted to maps from receivers
to sets of types, so the bottom element $\botcc$ maps every receiver to the set of all types,
and the top element $\topcc$ maps every receiver to the empty set of types.
During an actual execution, every receiver $r$ points to an object of some
runtime type. \otodo{What if $r$ points to null?} Therefore, a data flow fact
is unreachable along any feasible path if its corresponding lattice element
maps any receiver to the empty set of types.

@@

For each program point $n$, the result of an IFDS analysis gives
a set of facts $d$ that may be reached at $n$. The result of an IDE analysis
pairs each such fact~$d$ with a lattice element $\ell$. Formally,
for an IFDS problem $P$, the result $\result_{\text{IFDS}}$ has type
$N \to D$. Similarly, for an IFDS problem $Q$, the result $\result_{\text{IDE}}$
has type $N \to D \times L$.

@@

\otodo{The normalization function $\mathcal N$ was only needed in composition. I therefore
    ``inlined'' it into composition to save space. This makes composition a bit harder to intuitively
    verify, but saves complexity here. The verification can be in the proof in the supplementary appendix.}
The composition of two micro-function representations is defined as follows:
$\setpair{I}{U} \circ \setpair{I'}{U'} = \setpair{\lambda r\,.\,(I(r) \cap I'(r))\cup U(r)}{\lambda r\,.\,(I(r)\cap U'(r))\cup U(r)}$.
The following lemma shows that this implementation corresponds to the composition
of the denoted functions:
\begin{lemma}\label{lem:equal}
    For any pair of micro-function representations $\setpair{I}{U}$, $\setpair{I'}{U'}$,
    \[
        \denote{\setpair{I}{U}} \circ \denote{\setpair{I'}{U'}}
        = 
        \denote{\setpair{I}{U}} \circ \denote{\setpair{I'}{U'}}
    \]
\end{lemma}\mtodo{Should one of the sides of the equation be $\denote{\dots\circ\dots}$? Same for next lemma.}
\otodo{Prove this lemma in the appendix. The parts of the proof can be found in section 5.}
